---
title: "nic_demo"
output:
  pdf_document: default
  html_document: default
---

```{r}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```


# Introduction

My attempt to understand and reproduce the work done in the synthpop repo by j4sonli.

# Set up Environment

^ Only need to run once
```{r}
# 1. Initialise a renv in the project directory, you can check 
# it's created in your local machine by running `.libPaths()`
# renv::init()

# 2. Install the packages j4sonli recommended in his README
# renv::install(c(c("dplyr","tidyr","stringr","R.utils","tidycensus","bnlearn","foreach","doSNOW","ggplot2")))

# 3. Update your lockfile
# renv::snapshot()
```

# Load Libraries
```{r}
library(glue)
library(config)
library(dplyr)
library(ggplot2)
```


```{r}
# Test getting api key from config file
# Install if necessary: install.packages("config")
cfg <- config::get()
us_census_api_key <- cfg$us_census_api_key
```


```{r}
STATE_FIPS <- data.frame(
  state_name = c(
    "Alabama", "Alaska", "Arizona", "Arkansas", "California", "Colorado", 
    "Connecticut", "Delaware", "District of Columbia", "Florida", "Georgia", 
    "Hawaii", "Idaho", "Illinois", "Indiana", "Iowa", "Kansas", "Kentucky", 
    "Louisiana", "Maine", "Maryland", "Massachusetts", "Michigan", "Minnesota", 
    "Mississippi", "Missouri", "Montana", "Nebraska", "Nevada", "New Hampshire", 
    "New Jersey", "New Mexico", "New York", "North Carolina", "North Dakota", 
    "Ohio", "Oklahoma", "Oregon", "Pennsylvania", "Rhode Island", "South Carolina", 
    "South Dakota", "Tennessee", "Texas", "Utah", "Vermont", "Virginia", 
    "Washington", "West Virginia", "Wisconsin", "Wyoming"
  ),
  state_abb = c(
    "AL", "AK", "AZ", "AR", "CA", "CO", "CT", "DE", "DC", "FL", "GA", 
    "HI", "ID", "IL", "IN", "IA", "KS", "KY", "LA", "ME", "MD", "MA", 
    "MI", "MN", "MS", "MO", "MT", "NE", "NV", "NH", "NJ", "NM", "NY", 
    "NC", "ND", "OH", "OK", "OR", "PA", "RI", "SC", "SD", "TN", "TX", 
    "UT", "VT", "VA", "WA", "WV", "WI", "WY"
  ),
  state_code = c(
    "01", "02", "04", "05", "06", "08", "09", "10", "11", "12", "13", 
    "15", "16", "17", "18", "19", "20", "21", "22", "23", "24", "25", 
    "26", "27", "28", "29", "30", "31", "32", "33", "34", "35", "36", 
    "37", "38", "39", "40", "41", "42", "44", "45", "46", "47", "48", 
    "49", "50", "51", "53", "54", "55", "56"
  ),
  stringsAsFactors = FALSE
)

head(STATE_FIPS)
```


# Cambridge Massachusetts
Sample synth pop dataset in repo

```{r}
output_cambridge_hhs <- read.csv("./synthpop_output/2503302/syn_hhs_spatial_2503302-example.csv")

head(output_cambridge_hhs)
```


# Fulton County, Georgia

```{r include=FALSE}
# Subset to Georgia
STATE_NAME <- "Georgia"
COUNTY_FP <- 121 # Fulton county, where Georgia Tech is
puma_df <- read.csv("synthpop_data/2010_Census_Tract_to_2010_PUMA.csv")
GA_STATE_CODE <- STATE_FIPS[STATE_FIPS$state_name == STATE_NAME,]$state_code

GA_puma_df <- puma_df[puma_df$STATEFP == GA_STATE_CODE,]
GA_puma_codes <- (GA_puma_df %>%
  group_by(STATEFP, PUMA5CE) %>%
  summarise(count = n()))$PUMA5CE

# Print unique puma codes in GA, to run the synthpop script later
unique(GA_puma_codes)
```


```{r}
# Federal Information Processing Standard (Federal Information Processing Standard), two-digit numeric ids for US states
PUMA_CODE <- GA_puma_codes[1]
GA_STATE_CODE <- STATE_FIPS[STATE_FIPS$state_name == STATE_NAME,]$state_code
run_command <- glue('Rscript synthpop.R -puma={GA_STATE_CODE}00{PUMA_CODE}')

# Run the following command in your terminal to generate a synthpop of Fulton County, where Atlanta is a part of
print(run_command)
```
## Output

```{r}
# Where Atlanta is located
output_fulton_hhs <- read.csv("./synthpop_output/1300100/syn_hhs_1300100.csv")
head(output_fulton_hhs, 5)
```

##  Comparison (Cambridge v. Fulton)

```{r}
# Intersect columns from 
intersected_colnames <- intersect(colnames(output_cambridge_hhs), colnames(output_fulton_hhs))

for (colnames in intersected_colnames) {
  print(colnames)
}
```


```{r include=FALSE}
# Run line by line to see the unique output
unique(output_cambridge_hhs$tenur)
unique(output_cambridge_hhs$HHID)
unique(output_cambridge_hhs$tenur)
unique(output_cambridge_hhs$hhinc)
unique(output_cambridge_hhs$nwork)
unique(output_cambridge_hhs$ncars)
unique(output_cambridge_hhs$nchil)
unique(output_cambridge_hhs$hhsiz)
unique(output_cambridge_hhs$true_hhsiz)
unique(output_cambridge_hhs$true_nwork)
unique(output_cambridge_hhs$true_nchil)
unique(output_cambridge_hhs$hhtype)
unique(output_cambridge_hhs$puma)
unique(output_cambridge_hhs$i_age)
unique(output_cambridge_hhs$i_sex)
unique(output_cambridge_hhs$race_)
unique(output_cambridge_hhs$i_eth)
unique(output_cambridge_hhs$i_inc)
unique(output_cambridge_hhs$emply)
unique(output_cambridge_hhs$indus)
unique(output_cambridge_hhs$edu_q)
unique(output_cambridge_hhs$tenur_prox)
```

# Maps

##  Create directories

to store census information. Only run if you don't have the expected data directories.
```{r}
# base_path <- "synthpop_data/acs_marginals"
# 
# org_map_pumas <- c("0603701", "0603716", "0603733",
#                "4805000", "4804901", "4804614",
#                "0400134", "0400105", "0400122",
#                "2501400", "2503603", "2503302",
#                "1208900", "1203105", "1203103",
#                "5600200", "5600100", "5600300", "5600400", "5600500")
# 
# 
# for (map_puma in org_map_pumas) {
#   # Dynamically construct the path: "synthpop_data/acs_marginals/0603701"
#   full_puma_dir <- file.path(base_path, map_puma)
#   
#   if (!dir.exists(full_puma_dir)) {
#     dir.create(full_puma_dir, recursive = TRUE, showWarnings = FALSE)
#     cat("Initialized folder:", full_puma_dir, "\n")
#   }
# }
```

## Download Census Data For Demo
Only run again if you don't have the data yet.
```{r}
# org_map_pumas <- c("0603701", "0603716", "0603733",
#                "4805000", "4804901", "4804614",
#                "0400134", "0400105", "0400122",
#                "2501400", "2503603", "2503302",
#                "1208900", "1203105", "1203103",
#                "5600200", "5600100", "5600300", "5600400", "5600500")
# for (org_puma in org_map_pumas) {
#   command <- glue('Rscript synthpop_truncated.R -puma={org_puma}')
#   print(command)
# }
```


## Import `plot_maps`

```{r}
box::use(./plot_maps)
```


List of functions in plot_maps


* [X] plot_study_areas()
* [X] plot_error_heatmaps()
* [X] plot_error_map()
* [ ] plot_tract_hhtype_diffs_map()
* [ ] plot_pie()
* [X / Helper] error_heatmap()
* [Helper] rmse()
* [Helper] get_rmses()
* [Helper] process_marginals()
* [Helper] get_error_map_data()

## Constants
```{r}
map_pumas <- c("0603701", "0603716", "0603733",
               "4805000", "4804901", "4804614",
               "0400134", "0400105", "0400122",
               "2501400", "2503603", "2503302",
               "1208900", "1203105", "1203103",
               "5600200", "5600100", "5600300", "5600400", "5600500")

densities <- c(383.7, 9049.6, 42788.1,
               66.1, 3457.2, 11736.8,
               112.0, 3947.0, 8075.6,
               885.9, 5817.3, 28710.4,
               160.5, 1989.8, 3285.1,
               # last entry is density for entire state of WY
               5.97)
synpop_suffix <- "20220127"
```

# Trial Plot Functions

```{r}
# plot_study_areas()
box::reload(plot_maps)
plot_maps$plot_study_areas(map_pumas, save_pdf=FALSE)
```

```{r}
# plot_error_heatmap()
box::reload(plot_maps)
heatmap <- plot_maps$plot_error_heatmaps(syn_hhs_spatial_file="synthpop_output/2503302/syn_hhs_spatial_2503302-example.csv",
                          syn_indvs_spatial_file="synthpop_output/2503302/syn_indvs_spatial_2503302-example.csv",
                          marg_dir="synthpop_data/acs_marginals/2503302/", save_pdf=TRUE)
```

```{r}
# plot_error_map
map_pumas_subset <- c(2503302)
box::reload(plot_maps)
plot_maps$plot_error_map("Spatial errors in tenure by houshold income distribution",
               map_pumas_subset, densities,
               variable="tenur_hhinc",
               spatial_level="tract",
               synpop_suffix=synpop_suffix,
               gray_hh_pop_under=30,
               gray_indv_pop_under=50,
               NA_label="too few households",
               save_pdf=TRUE)
#synthpop/synthpop_output/2503302/syn_hhs_spatial_2503302.csv
```



# tmap fixes

## tmap
```{r}
# .libPaths()
# # Re-install the v3 branch directly to renv environment
# renv::purge("tmap")
# renv::install("r-tmap/tmap@v3")
# 
# # Snapshot the environment state to update your renv.lock file
# renv::snapshot(prompt = FALSE)
# packageVersion("tmaptools")
```

## tmaptools
```{r}
# # 1. Wipe the current mismatched tmaptools
# renv::purge("tmaptools")
# 
# # 2. Install the legacy version of tmaptools that matches tmap v3
# # Version 3.1-1 is the stable companion release for tmap v3
# renv::install("cran/tmaptools@3.1-1")
# 
# # 3. Reinstall tmap@v3 just to ensure dependencies hook back together cleanly
# renv::install("r-tmap/tmap@v3", upgrade = "never")
# 
# # 4. Lock this working combination into your renv lockfile
# renv::snapshot(prompt = FALSE)
```






